---
title: "Lab 8 - Linear Regression in R"
author: "Daphne Virlar"
date: "11/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Load libraries and data
```{r}
library(tidyverse)
salmon <- read_csv("salmon_imports_copy.csv")
```


Exploratory graphs
```{r}
expo_salmon <- ggplot(salmon, aes(x=year, y= salmon_fresh))+
  geom_point()

expo_salmon

#It looks mostly fine, except for those two weird dips
```



Create the linear regression
```{r}
#Outline for a linear regression
      #model_name <- lm(y_variable ~ x_variable, data = df_name)
salmon_model <- lm(salmon_fresh ~ year, data = salmon)
summary(salmon_model)
  #intercept --> y-intercept
  #year      --> slope
  
#R squared value = 0.9478
#Adjusted = 0.9458       --> use adjusted r square variable when using multiple predictor variables


#y = β0 + β1x + e
      #y = (linear component) + error term

b <- -14982940
m <- 7550

#Write out the linaer model
#salmon_imports <- m*x + b
#y = 7550x + (-14982940)

```


What does the slope mean in the context of this model?
It means that salmon imports are increasing per year.

What about the y-intercept? Why is that concerning? What does this mean about extrapolating this model for past values?
It's concerning because you can't have less than zero salmon. It makes it impossible because at year zero, we have "negative salmon", and you can't have negative fish in the ocean. 



Model Diagnostics
```{r}
plot(salmon_model)
#Our residuals look fine. Normal Q-Q and Residuals v Fitted have some outliers, but I'm not concerned about it given what our exploratory data analysis shows us. 



```

Explore model fit and significance
```{r}
summary(salmon_model)
#H0 -- no significant correlation y and x
#HA -- there is significant correlation between y and x

#Yes, year significantly predicts salmon imports. R2 = 0.9478, p-value < 0.001


#The R^2 value tells us how much variance in our dependent variable is expained by our independent variable


#About 5% of salmon imports is not explained by year

```


Prepare a final graph
```{r}
#create a final graph of the linear regression model with the original data, model and 95% confidence interval for predictions


salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh))+
  geom_point()+
  geom_smooth(method = lm, se = TRUE, size = 0.5, color = "purple") + 
  #method = lm tells R to do this line as a linear regression
#se = TRUE tells R to include a confidence interval
#size = 0.5 tells r to change the width of the line
#color changes the color of the line
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017))+
  labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports \n(1,000 lbs)")+
  ggtitle("Salmon Imports by Year")


salmon_final_graph
```



Make predictions using the model
```{r}
#Column name is same as year that the model uses. This is intentional
new_years <- data.frame(year = c(2022, 2024, 2026, 2028))


future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")
future_predict
  #fit is the model fit
  #lwr is the lower bound, and vice versa for upr

predictions <- data.frame(new_years, future_predict)


```


Find Pearson's r (correlation)
```{r}
#Use the cor.test() function to find Pearson’s r for the linear relationship described between year and salmon imports.

#cor.test(x,y)
salmon_pr <- cor.test(salmon$year, salmon$salmon_fresh)
salmon_pr


#Pearson's r = 0.9735387
#p-value < 0.001

#Strong positive correlation

```

Describe Results


F-statistic: 471.9 on 1 and 26 DF,  p-value: < 2.2e-16


1. Thorough
“Year significantly predicts salmon imports (b = 7550, t(df) = , p < 0.01) with a strong positive correlation between the two (Pearson's r = 0.97). The overall model (salmon imports = 7550*year - 14982940, in tonnes) explains a significant amount of variance in salmon imports (F(1,26) = 471.9, p < 0.01, R2 = 0.95)



2. Simplified
“The model (salmon imports = 7550 ∗ (year) − 14982940; in tonnes) explains a significant amount of
variance in salmon imports (F(1,26) = 471.9, p < 0.001, R2 = 0.95, Pearson’s r = 0.97).”






